\huge UNet-HCRF Integration for EEG Pattern Recognition

### 2024-12-12 manuscript/main.tex
edit intro

\section{Introduction}
%Introduction: problem definition and motivation.

\sx{EEG tasks}

Sequence labeling tasks are pivotal in diverse domains, especially healthcare,
where precise and timely data classification can significantly impact patient
outcomes. Medical datasets, such as physiological signals, often exhibit
complexity and heterogeneity, posing challenges for traditional modeling
techniques. Therefore, developing models capable of extracting meaningful
patterns from these datasets is crucial for accurate diagnostic and prognostic
decisions.

In recent years, deep learning has emerged as a powerful tool for sequence
labeling. Neural network architectures like long short-term memory (LSTM) and
Transformer models have revolutionized sequential data processing.
Additionally, conditional random fields (CRFs) have been employed for modeling
label dependencies, demonstrating promising results.
However, while these methods have shown individual strengths, a comprehensive
approach that leverages their complementary capabilities is essential for
tackling complex, multi-modal datasets.

This project introduces a novel framework that combines deep neural networks
and probabilistic graphical models (PGMs) to enhance sequence labeling
performance. By employing Transformer~\citep{vaswani2017attention} for feature
extraction and integrating them with CRF models,
we aim to improve sequence tagging accuracy. Furthermore, mean-field
approximation techniques are utilized for efficient inference, ensuring a
computationally efficient yet effective approach.
This integration of modern neural network architectures with CRFs offers
a robust and promising solution for complex sequence labeling tasks.

>>>>

\section{Introduction}

Electroencephalography (EEG) pattern recognition plays a pivotal role in
various fields, including neuroscience, clinical medicine, and human-computer
interaction. By analyzing brain electrical activity, we can gain deeper
insights into brain states and functions, providing a scientific basis for
diagnosing and treating neurological disorders, as well as developing
brain-computer interfaces. EEG pattern recognition not only helps to unveil the
mechanisms of the brain but also enables the development of smarter and more
human-centric technologies. However, traditional manual analysis of EEG signals
is a time-consuming and subjective process, often leading to inconsistencies in
diagnosis~\citep{liu2013recognizing}.

In recent years, deep learning has emerged as a powerful tool to automate and
improve EEG pattern recognition~\citep{craik2019deep}. Neural network
architectures like convolutional neural network (CNN), long short-term memory
(LSTM) and Transformer models have revolutionized EEG data processing.
Additionally, hidden conditional random fields (HCRFs) have been employed for
modeling label dependencies, demonstrating promising results.
While these methods have shown individual strengths, a comprehensive
approach that leverages their complementary capabilities is essential for
tackling complex, multi-modal datasets.

This project introduces a novel framework that combines deep neural networks
and probabilistic graphical models (PGMs) to enhance the accuracy of EEG
pattern recognition . By employing UNet~\citep{ronneberger2015unet} for feature
extraction and integrating them with HCRF models, we aim to improve the
precision of EEG pattern recognition tagging. Furthermore, mean-field
approximation techniques are utilized for efficient inference, ensuring a
computationally efficient yet effective approach.
This integration of modern neural network architectures with HCRFs offers
a robust and promising solution for EEG pattern recognition tasks.
